Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-01 paper/paper_new.tex

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digital specimen accessible through global, open access repositories; they enable researchers, educators, students, and the general public to examine and compare in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these data items, due to both image quality and metadata issues.
%{jane}should we try to insert machine learning here, last sentence, or computation methods ok?

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images, e.g. with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements, provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of 7,247
images drawn from the Illinois Natural History Survey (INHS) Collection.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms \cite{dc_terms}, the Ecological Metadata Language (EML) \cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.

@@

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the mean foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.9. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.

>>>>

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which contain specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digital specimen images are accessible through global, open access repositories; they enable researchers, educators, students, and the general public to examine and compare collections in ways previously unimaginable. Significant challenges surface, however, when researchers seek to apply computational methods to these repositories, due to both image quality and metadata issues.
%{jane}should we try to insert machine learning here, last sentence, or computation methods ok?

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images---e.g.\ those with low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements---provide inadequate input
for automated
image analysis and machine learning algorithms, and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/cm) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images
to determine if, for instance, a specific fish grows to different lengths
in different habitats.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

Since the digital collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.
As a step towards improving metadata in specimen research image collections, members of Drexel University's Metadata Research Center are developing methods to automatically analyze fish images and
extract a set of data features that provide important metadata about the
digitized specimen.
The research is being conducted as part of the Biology Guided Neural Networks (BGNN) project, which aims to develop a novel class of artificial neural networks that exploit machine readable and predictive knowledge associated with specimen images, phylogenies and anatomy ontologies.
Using a combination of machine learning and image informatics tools and techniques, we can accurately determine general image quality and metadata, such as fish quantity and location within images, fish orientation, and
image scaling based on ruler identification and measurement. Image scaling
allows us to compute quantitative features about the fish specimens, such
as their length and area.  In order to test and validate our methods, they
were  applied to a set of \(7,244\)
images drawn from the Illinois Natural History Survey (INHS) Collection.
Figure \ref{fig:INHS_examples} presents three typical images used in our
study.
The following section of the paper provides contextual background for this work, followed by the research goals and objectives, and a review of our research methods. Next, the results are presented, followed by a discussion. The conclusion highlights key findings and identifies next steps.

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. Darwin Core~\cite{biodiv_info_standards} developed specifically for digital museum projects is one of the most frequently used standards. Another approach is to create a metadata application profile by integrating properties from Dublin Core's DCMI Metadata Terms~\cite{dc_terms}, the Ecological Metadata Language (EML)~\cite{EML_2019} and other namespaces. There is also the recent Audubon Core supporting the discoverability, dissemination, and use of data relating to biological organisms, including digitized specimens.~\cite{audub_core} These descriptive standards include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is limited, particularly for researchers seeking to apply machine learning, and leverage both the image and the metadata for scientific research.

@@

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the mean foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.6. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.
